A Study of Immersion Factors Using an Augmented Reality Simulation

### 2009-06-24 trunk/Paper/paper.tex
working on analysis section

Livingston and Ai \cite{4637329} studied the effects of various sources of registration error with an AR simulation similar to our own.  Here, participants tracked a virtual car moving throughout a real environment, with a white box representing an augmented view of the car's location.  A white box was continuously visible, even when the car itself was occluded by a building in the environment.  Other virtual cars and their associated white boxes acted as distractors.  At specific times during the experiment, the simulation would freeze and the participant was asked to align the center of their view (indicated by cross-hairs) on the location of the correct car and press a key.  While the effects of several types of error were studied in this work, the effect of registration dropouts was not examined.

Bane and Hollerer \cite{1383060} propose a set of AR tools for virtual x-ray vision.  Various augmented views are presented, each attempting to overcome the "Superman's X-ray vision" problem of presenting too much AR information to the user.  The toolset's usefulness is illustrated with an example of an outdoors user viewing the contents of a nearby building.

other studies which vary AR fov, tracking failures

previous X-ray AR work

simulated AR

\cite{4637329,4811058,1383060}

@@

\begin{figure}[ht!]
	\centering
	\includegraphics[width=3in]{figures/augmentedroom.png}
	\caption{Overhead view of virtual room (approx. measure).  Participant is stationary in the center of the room, with initial orientation along ``user view'' arrow.  Top-hat's initial position is outside the front doorway, directly in the participant's view.}
\end{figure}

@@

During each trial there are 20 people in addition to a virtual man \emph{Tophat} wearing a tall black tophat walking an unpredictable path outside of the room.  The paths for each virtual person was randomly generated from set of coordinates exterior to the room.  Each path had the constraint that they must stay within the rectangular $50\times50$ $m^2$ area surrounding the room.  At each path point, the person will randomly change his/her walking speed within the range of 4 to 7 m/s.  \emph{Tophat} had the additional constraints as follows.  He must start at the same location for each trial which made him visible by the participant through the front door the building.  In order to make the paths by of similar difficulty \emph{Tophat} must walk around the entire building at least once, and he must be visible through the windows for at least 10 seconds.  These constraints were met by randomly generating paths until all were satisfied.  We generated 27 sets of paths in this fashion, each corresponding to a specific trial.  We used a latin squares ordering to discourage the effects of learning.

@@

@@

We hypothesize that less immersive conditions will have longer and more frequent periods where the participant is lost.  We considered several different metrics which might be appropriate in testing this hypothesis.

@@

We can more explicitly detect the lost state by applying a threshold $\tau$ to the data.  When the error rises above the threshold, we consider the participant lost.  Here, we use a 12 degree error as the threshold.

One metric which uses this threshold is the ``time to failure,'' the time until a lost state is encountered.  However a participant may reach the lost state sooner or later depending on the movement of the avatars.  We also considered a measurement we called ``time tracking,'' which is the amount of time spent in the tracking state.  This metric seems to most clearly represent how well a participant performed.  Trials with either more lost periods or longer lost periods will result in a lower total time tracking.

\begin{figure}[ht!]
	\centering
	\includegraphics[width=3in]{figures/tophatscreenshot.png}
	\caption{Sample user-view of AR simulation.  \emph{Tophat} is visible (non-occluded) in the doorway.  The augmented view overlays transparent red rectangles on the characters, indicating their location while occluded.}
\end{figure}

@@

>>>>

Livingston and Ai studied the effects of various sources of registration error with an AR simulation similar to our own \cite{4637329}.  Here, participants tracked a virtual car moving throughout a real environment, with a white box representing an augmented view of the car's location.  A white box was continuously visible, even when the car itself was occluded by a building in the environment.  Other virtual cars and their associated white boxes acted as distractors.  At specific times during the experiment, the simulation would freeze and the participant was asked to align the center of their view (indicated by cross-hairs) on the location of the correct car and press a key.  While the effects of several types of error were studied in this work, the effect of registration dropouts was not examined.

Bane and H\"{o}llerer propose a set of AR tools for virtual x-ray vision \cite{1383060}.  Various augmented views are presented, each attempting to overcome the "Superman's X-ray vision" problem of presenting too much AR information to the user.  The toolset's usefulness is illustrated with an example of an outdoors user viewing the contents of a nearby building.

% TODO: add discussion of validity of AR simulation w/ citations

%other studies which vary AR fov, tracking failures?

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=3in]{figures/augmentedroom.png}
	\caption{Overhead view of virtual room (approx. measure).  Participant is stationary in the center of the room, with initial orientation along ``user view'' arrow.  Top-hat's initial position is outside the front doorway, directly in the participant's view.}
\end{figure}

@@

During each trial there are 20 people in addition to a virtual man \emph{Top Hat} wearing a tall black top hat walking an unpredictable path outside of the room.  The paths for each virtual person was randomly generated from set of coordinates exterior to the room.  Each path had the constraint that they must stay within the rectangular $50\times50$ $m^2$ area surrounding the room.  At each path point, the person will randomly change his/her walking speed within the range of 4 to 7 m/s.  \emph{Top Hat} had the additional constraints as follows.  He must start at the same location for each trial which made him visible by the participant through the front door the building.  In order to make the paths by of similar difficulty \emph{Tophat} must walk around the entire building at least once, and he must be visible through the windows for at least 10 seconds.  These constraints were met by randomly generating paths until all were satisfied.  We generated 27 sets of paths in this fashion, each corresponding to a specific trial.  We used a latin squares ordering to discourage the effects of learning.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=3in]{figures/tophatscreenshot.png}
	\caption{Sample user-view of AR simulation.  \emph{Top Hat} is visible (non-occluded) in the doorway.  The augmented view overlays transparent red rectangles on the characters, indicating their location while occluded.}
\end{figure}

@@

% TODO: include two plots of error here (one with losses, one without)

We hypothesize that less immersive conditions will have longer and more frequent periods where the participant loses the target.  We considered several different metrics which might be appropriate in testing this hypothesis.

@@

% median ??

We can more explicitly detect the participant's state by applying a threshold $\tau$ to the data.  The threshold specifies the maximum angular error that still represents successful tracking of the target. 
%When the error rises above the threshold, we consider the participant lost.
We use a threshold of one quarter of the total horizontal field of view of the HMD, $\tau = 12$ degrees.
% TODO: how to motivate this threshold?

Using this threshold we measure a participant's ``time to failure,'' the time until the a lost state is encountered.  However, this may not be a very descriptive measure, since a participant may reach the lost state sooner or later depending on the movement of the avatars, which varies between trials.

We also consider a measurement we call ``time tracking,'' which is the amount of time spent in the tracking state.  This metric seems to most generally represent how well a participant performed.  Trials with either more lost periods or longer lost periods will result in a lower total time tracking.

For a more specific measure of performance, we count the number of times a participant reaches the lost state in a trial.  We only record a lost state when the error stays above the threshold $\tau$ for a minimum of 0.5 seconds.
% TODO: how to motivate this threshold?

%time tracking is overview metric -- captures both number of times lost and length of lost periods

%number of times lost is piece of overview -- number of times lost.

@@

%\begin{figure}[t]
%	\centering
%	\includegraphics[width=3in]{figures/numtimes_deadlen.pdf}
%	\caption{Box plot of the number of times the participant is lost versus the length of the tracking failure periods.}
%\end{figure}

%\begin{figure}[t]
%	\centering
%	\includegraphics[width=3in]{figures/numtimes_fov.pdf}
%	\caption{Box plot of the number of times the participant becomes lost versus the field of view of the augmentations.}
%\end{figure}

%\begin{figure}[t]
%	\centering
%	\includegraphics[width=3in]{figures/numtimes_interaction.pdf}
%	\caption{Interaction plot of the effect of field of view and tracking failure on the number of times the participant becomes lost.}
%\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=3in]{figures/tt_deadlen.pdf}
	\caption{\label{fig:tt_deadlen}Box plot of the amount of time that the target is near the crosshairs versus the length of tracker failure periods.}
\end{figure}

%\begin{figure}[t]
%	\centering
%	\includegraphics[width=3in]{figures/tt_fov.pdf}
%	\caption{Box plot of time the participant is not lost versus the field of view of the augmentations.}
%\end{figure}

%\begin{figure}[t]
%	\centering
%	\includegraphics[width=3in]{figures/numtimes_extra_deadlen.pdf}
%	\caption{Box plot of the number of times the participant is lost versus the length of the tracking failure periods, for the medium field of view condition, including the half second trial.}
%\end{figure}

% TODO: should we include data from older generation??

% TODO: look at analysis from mean.

We performed two-way {ANOVA} for the time tracking metric as an overall analysis of the effects of our two factors.  Table \ref{fig:anova} gives the results of this analysis.  Both the failure length and the field of view had a very significant effect on performance using this metric.  
% TODO: interaction?

% TODO: num times lost: fov not significant, deadlen significant, interaction -- talk about what interaction means.

Figure \ref{fig:tt_deadlen} shows a box plot of time tracking versus failure length.  Performance decreases between zero and one seconds, but seems to level off between one and two.  We ran a post-hoc analysis to determine which conditions were significantly different.  Using Tukey's HSD, we found that conditions one and two were significantly different than zero (p<??), but not significantly different from each other (p<??).  This suggests that we may be seeing a thresholding effect above one second tracking failures.

This suggests that we have reached a threshold at one second.  We had seven participants complete an extra three trials with the medium field of view and tracking failure periods of 0.5 seconds.  Figure 8 shows the data.  There wasn't a significant difference between the zero length and half second length conditions.  For future work, we could find the threshold between 0.5 and 1.

%use multiple comparison of means to determine which pairs are significant -- difference between length of one and length of two is not significant.

\begin{figure}[htb]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
%\multicolumn{2}{|c|}{Selection Stroke Thresholds} \\
Metric & Factor & F value & Pr \\
\hline
Time Tracking & Failure Length & 20.41 & $<0.001$\\
Time Tracking & Field of View & 13.549 & $<0.001$\\
Number of Times Lost & Failure Length & 61.816 & $<0.001$\\
\hline
\end{tabular}
 \caption{\label{fig:anova}Factors which have a significant effect.}
\end{figure}
